---
title: "Hands-on Exercise 2"
author: "Ten Wei Peng"
date: "April 20, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---



## Content

::: {style="font-size: 1.5em"}
-   Beyond ggplot2 Themes

-   Beyond ggplot2 Annotation

-   Beyond ggplot2 facet
:::

## Getting started

### Installing and loading the required libraries

::: columns
::: {.column width="50%"}
-   Before we get started, it is important for us to ensure that the required R packages have been installed. If yes, we will load the R packages. If they have yet to be installed, we will install the R packages and load them onto R environment.

-   The chunk code on the right will do the trick.
:::

::: {.column width="50%"}
::: {style="font-size: 1.2em"}
```{r}
pacman::p_load(tidyverse, patchwork, 
               ggthemes, hrbrthemes,
               ggrepel) 
```
:::
:::
:::

------------------------------------------------------------------------

### Importing data

::: columns
::: {.column width="50%"}
-   The code chunk below imports *exam_data.csv* into R environment using [*read_csv()*](https://readr.tidyverse.org/reference/read_delim.html) function of [**readr**](https://readr.tidyverse.org/) package.

-   **readr** is one of the tidyverse package.

::: {style="font-size: 1.2em"}
```{r}
exam_data <- read_csv("data/Exam_data.csv")
```
:::
:::

::: {.column width="50%"}
-   Year end examination grades of a cohort of primary 3 students from a local school.

-   There are a total of seven attributes. Four of them are categorical data type and the other three are in continuous data type.

    -   The categorical attributes are: ID, CLASS, GENDER and RACE.
    -   The continuous attributes are: MATHS, ENGLISH and SCIENCE.
:::
:::

## Beyond ggplot2 Annotation

::: columns
::: {.column width="50%"}
One of the challenge in plotting statistical graph is annotation, especially with large number of data points.

::: {style="font-size: 1.2em"}
```{r}
#| eval: false
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
:::
:::

::: {.column width="50%"}
```{r}
#| echo: false
#| fig-height: 9
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
:::
:::

------------------------------------------------------------------------

### Working with ggrepel

::: columns
::: {.column width="50%"}
[**ggrepel**](https://ggrepel.slowkow.com/index.html) is an extension of **ggplot2** package which provides `geoms` for **ggplot2** to repel overlapping text as in our examples on the right. We simply replace `geom_text()` by [`geom_text_repel()`](https://ggrepel.slowkow.com/reference/geom_text_repel.html) and `geom_label()` by [`geom_label_repel`](https://ggrepel.slowkow.com/reference/geom_text_repel.html).

::: {style="font-size: 1.2em"}
```{r}
#| eval: false
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
:::
:::

::: {.column width="50%"}
```{r}
#| echo: false
#| fig-height: 9
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
:::
:::

## Beyond ggplot2 Themes

::: columns
::: {.column width="50%"}
ggplot2 comes with eight [built-in themes](https://ggplot2.tidyverse.org/reference/ggtheme.html), they are: `theme_gray()`, `theme_bw()`, `theme_classic()`, `theme_dark()`, `theme_light()`, `theme_linedraw()`, `theme_minimal()`, and `theme_void()`.

::: {style="font-size: 1.2em"}
```{r}
#| eval: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +  
  ggtitle("Distribution of Maths scores") 

```
:::

::: {style="font-size: 0.65em"}
Refer to this [link](https://ggplot2.tidyverse.org/reference/index.html#themes) to learn more about ggplot2 `Themes`
:::
:::

::: {.column width="50%"}
```{r}
#| echo: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 

```
:::
:::

------------------------------------------------------------------------

### Working with ggtheme package

::: columns
::: {.column width="50%"}
[**ggthemes**](https://cran.r-project.org/web/packages/ggthemes/index.html) provides ['ggplot2' themes](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/) that replicate the look of plots by Edward Tufte, Stephen Few, [Fivethirtyeight](https://fivethirtyeight.com/), [The Economist](https://www.economist.com/graphic-detail), 'Stata', 'Excel', and [The Wall Street Journal](https://www.pinterest.com/wsjgraphics/wsj-graphics/), among others.

::: {style="font-size: 1.2em"}
```{r}
#| eval: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```
:::

-   It also provides some extra geoms and scales for 'ggplot2'. Consult [this vignette](https://mran.microsoft.com/snapshot/2016-12-28/web/packages/ggthemes/vignettes/ggthemes.html) to learn more.
:::

::: {.column width="50%"}
```{r echo=FALSE}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```
:::
:::

------------------------------------------------------------------------

### Working with hrbthems package

::: columns
::: {.column width="50%"}
[**hrbrthemes**](https://cinc.rud.is/web/packages/hrbrthemes/) package provides a base theme that focuses on typographic elements, including where various labels are placed as well as the fonts that are used.

::: {style="font-size: 1.2em"}
```{r}
#| eval: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```
:::

-   The second goal centers around productivity for a production workflow. In fact, this "production workflow" is the context for where the elements of hrbrthemes should be used. Consult [this vignette](https://cran.r-project.org/web/packages/hrbrthemes/vignettes/why_hrbrthemes.html) to learn more.
:::

::: {.column width="50%"}
```{r}
#| echo: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```
:::
:::