---
title: "In-class Exercise 02"
author: "Ten Wei Peng"
date: "April 20, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Loading Packages

::: callout-note
In the code chunk, p_load() and [**pacman**](https://cran.r-project.org/web/packages/pacman/index.html) package is used to load tidyverse family of packages
:::

```{r}
pacman::p_load(tidyverse, ggdist,ggthemes, colorspace, ggridges)
```

## Importing Data

```{r}
exam_df <-read_csv("data/Exam_data.csv")
```

# Visualisation Distribution

## Histogram
::: panel-tabset

### The Task
Using the step you learned , build a histogram


### The Code

```{r}
ggplot(data=exam_df, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of English scores") +
  theme_economist()
```

:::



## Distribution
::: panel-tabset

### The Task
Using the step you learned , build a density plot


### The Code

```{r}
ggplot(data=exam_df, 
             aes(x = ENGLISH)) +
  geom_density(color ="#1696d2",
               adjust =0.65,
               alpha =0.6)
```

:::






## The alternative design
::: panel-tabset

### The Task
Using the step you learned , build a density plot


### The Code

```{r}
median_eng <- median(exam_df$ENGLISH)
mean_eng <- mean(exam_df$ENGLISH)
std_eng <- sd(exam_df$ENGLISH)

ggplot(data = exam_df, aes(x = ENGLISH)) +
  geom_density(color = "#1696d2", adjust = 0.65, alpha = 0.6) +
  stat_function(
    fun = dnorm,
    args = list(mean = mean_eng, sd = std_eng),
    col = "grey30",
    size = 0.8
  ) +
  geom_vline(
    aes(xintercept = mean_eng),
    colour = "#4d5887",
    linewidth = 0.6,
    linetype = "dashed"
  ) +
  annotate(
    geom = "text",
    x = mean_eng - 8,
    y = 0.04,
    label = paste0("Mean ENGLISH: ", round(mean_eng, 2)),
    color = "#4d5887"
  ) +
  geom_vline(
    aes(xintercept = median_eng),
    colour = "#00AFBB",
    linewidth = 0.6,
    linetype = "dashed"
  ) +
  annotate(
    geom = "text",
    x = mean_eng - 8,
    y = 0.04,
    label = paste0("Median ENGLISH: ", round(median_eng, 2)),
    color = "#00AFBB"
  )
```

:::







