---
title: "Hands-on Exercise 1"
---

## Getting Started

### Installing and loading the required libraries

::: callout-note
The code chunk on the right assumes that you already have [**pacman**](https://cran.r-project.org/web/packages/pacman/index.html) package installed. If not, please go ahead install pacman first.
:::

::: {.column width="50%"}
```{r}
pacman::p_load(tidyverse)
```
:::

### Importing data

:::
-   The code chunk below imports *exam_data.csv* into R environment by using [*read_csv()*](https://readr.tidyverse.org/reference/read_delim.html) function of [**readr**](https://readr.tidyverse.org/) package.

-   **readr** is one of the tidyverse package.

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

-   Year end examination grades of a cohort of primary 3 students from a local school.

-   There are a total of seven attributes. Four of them are categorical data type and the other three are in continuous data type.

    -   The categorical attributes are: ID, CLASS, GENDER and RACE.
    -   The continuous attributes are: MATHS, ENGLISH and SCIENCE.
:::












